---
title: "For Loop practice"
author: "Allie Cole"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

### Code warm up 

```{r}

quater_splits <- c(1, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

vec_run <- vector(mode = "numeric", length = length(quater_splits) - 1)

for (i in seq_along(quater_splits) - 1) {
  p <- (quater_splits[i] + quater_splits[i +1])
  vec_run[i] <- p
}

```

### Writing some functions, basics

```{r}

birddog_sum <- function(bird, dog) { #needs these two arguements to work 
  add_birddog <- bird + dog  #this is what teh function will do 
  return(add_birddog) #this is what it will return 
}

birddog_sum(bird = 100, dog = 5)

birddog_sum(bird = "hawk", dog = 10) #cant do this! r will give you an error
```

### adding messages and error notifications

```{r}
birddog_sum <- function(bird, dog) { 
  
  if (bird > 50) {
    stop("Birds cant be greater than 50!") #this is an error and will stop your code from running 
  }
  
  if(bird <10 ) {
    message("cool little bird")  #this is just a message 
  }
  
  if(dog <0) {
    warning("Can a dog be less than zero?")  #this si a warning 
  }
  
  add_birddog <- bird + dog   
  return(add_birddog)  
}

birddog_sum(bird = 7, dog = 5)

birddog_sum(bird = "hawk", dog = 10)
```



### function with strings 


```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello,", friend, "have a great", weekday))
}

say_hello(friend = "Sarah", weekday = "Thursday")

```


### Print v return example 

```{r}

find_multiple <- function(a, b) {
  my_mult <- a * b 
  print(my_mult)
}

find_multiple(a= 3, b = 4)



```

### functions with conditions 

what you want the function to do can change based on what you are putting into it 

```{r}
animal_age <- function(animal, age) {
  
  if (!animal %in% c("dog", "goat")) {  #using the in operator! 
    warning("Animal must be a dog or a goat")
  }
  
  if (animal == "dog") {
    creature_age <- age * 7
  }
  
  else if (animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

animal_age(animal = "dog", age = 4)

# animal_age(animal = "hawk", age =  3) #will break! also if you put numbers 
```

### looking at %in%


```{r}
adelie_gentoo <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo")) # look in every row in species colum does it match A or G 
#this will look in any row for Adelie or Gentoo 
```


```{r}
adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))

# when I say look for an exact match with a vector, it is looking for an exact match with the order Adelie and then Gentoo. if not it will be rejected 

# could do this, 

adelie_gentoo_wrong <- penguins %>% 
  filter(species == "Adelie" | species == "Gentoo") # but this will take forever if you have lots of stuff
```

### functions with for loops 

```{r}
df_means <- function(df) {
  
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE) 
    col_name <- colnames(df[i])
    print(paste("The mean value that", col_name, "is", col_mean))
  }
}

df_means(df = mtcars)

# upate the function so that for each iteration is returns the phrase "The mean value that column name 
```

### making a subset of only teh structural size variablles from teh penguins data set for chinstrap

```{r}
chinstrap_sizes <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, flipper_length_mm, bill_depth_mm, body_mass_g)

#now lets use our new function

df_means(df = chinstrap_sizes)
```

### condtion to check teh class of an input 

```{r}
lunchtime <- function(tacos, price_per) {
  
  if (is.numeric(tacos) == FALSE) {
    stop("you must input a number of tacos")
  }
  
  total_price <- tacos * price_per
  return(total_price)
}

lunchtime(tacos = 4, price_per =  2.5)

# lunchtime(tacos = "hi", price_per =  2.5)
```

### Use values from a data frame to calculte something based on user input 

```{r}
# calculate the ground water transpot rate with the filter function


gw_rate <- function(site){
  
  gw_slopes_depths <- data.frame(sitename = c("Mountian", "Prarie", "Desert", "Beach"), 
                                 depth = c(32, 41, 63, 2), 
                                 slope = c(11.2, 0.4, 0.8, 2.6))
  
  site_select <- filter(gw_slopes_depths, sitename == site)
  
  transport_rate <- 1.4 * site_select$slope + 3.6 * site_select$depth
  
  return(transport-rate)
}

gw_rate(site = "Prarie")


```


```{r}
# now we are going to do the exact same thing but use switch 

gw_rate_switch <- function(site) {
  
  gw_slopes_depths <- switch(site, 
                             "Mountian" = c(32, 11.2), 
                             "Prarie" = c(41, 0.4), 
                             "Desert" = c(63, 0.8), 
                             "Beach" = c(2, 2.6))
  transport_rate <- 1.4 * gw_slopes_depths[2] + 3.6 * gw_slopes_depths[1]
  return(transport_rate)
}

gw_rate_switch(site = "Prarie")

```

```{r}
  
logistic_growth <- function(K, N0, r, time) {
  
  Nt <- K / (1 + ((K - N0)/ N0) * exp(-r * time))
  print(Nt)
}

logistic_growth(K = 5000, N0 = 200, r = 0.43, time = 6)

time_vec <- seq(from = 0, to = 35, by = 0.1)

bac_growth <- logistic_growth(K = 5000, N0 = 200, r = 0.43, time = time_vec)

bac_df <- data.frame(time_vec, bac_growth)


```


```{r}
ggplot(data = bac_df, aes(x = time_vec, y = bac_growth)) +
  geom_jitter()
```


```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1) #same as above 

out_matrix <- matrix(ncol = length(r_seq), nrow = length(time_vec))

# N0 = 100 K = 6000

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 600, N0 = 100, r = r_seq[i], time = time_vec[j])
    out_matrix[j, i] <- pop
  }
}
 out_matrix
```

### now a little wrangling

```{r}
# attach the time vector as teh first column of our out_matrix 

out_df <- data.frame(time = time_vec, out_matrix)

#give columns good names 

colnames(out_df) <- c("time", paste0("gr_", r_seq))

#make it long form (or tidy version) for using pivot longer 

out_df_long <- out_df %>% 
  pivot_longer(-time, names_to = "growth_rate", values_to = "population")
```

### Plotting it!


```{r}
ggplot(data = out_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate))
```










