---
title: "For Loop practice"
author: "Allie Cole"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

### Code warm up 

```{r}

quater_splits <- c(1, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

vec_run <- vector(mode = "numeric", length = length(quater_splits) - 1)

for (i in seq_along(quater_splits) - 1) {
  p <- (quater_splits[i] + quater_splits[i +1])
  vec_run[i] <- p
}

```

### Writing some functions, basics

```{r}

birddog_sum <- function(bird, dog) { #needs these two arguements to work 
  add_birddog <- bird + dog  #this is what teh function will do 
  return(add_birddog) #this is what it will return 
}

birddog_sum(bird = 100, dog = 5)

birddog_sum(bird = "hawk", dog = 10) #cant do this! r will give you an error
```

### adding messages and error notifications

```{r}
birddog_sum <- function(bird, dog) { 
  
  if (bird > 50) {
    stop("Birds cant be greater than 50!") #this is an error and will stop your code from running 
  }
  
  if(bird <10 ) {
    message("cool little bird")  #this is just a message 
  }
  
  if(dog <0) {
    warning("Can a dog be less than zero?")  #this si a warning 
  }
  
  add_birddog <- bird + dog   
  return(add_birddog)  
}

birddog_sum(bird = 7, dog = 5)

birddog_sum(bird = "hawk", dog = 10)
```



### function with strings 


```{r}
say_hello <- function(friend, weekday) {
  print(paste("Hello,", friend, "have a great", weekday))
}

say_hello(friend = "Sarah", weekday = "Thursday")

```


### Print v return example 

```{r}

find_multiple <- function(a, b) {
  my_mult <- a * b 
  print(my_mult)
}

find_multiple(a= 3, b = 4)



```

### functions with conditions 

what you want the function to do can change based on what you are putting into it 

```{r}
animal_age <- function(animal, age) {
  
  if (!animal %in% c("dog", "goat")) {  #using the in operator! 
    warning("Animal must be a dog or a goat")
  }
  
  if (animal == "dog") {
    creature_age <- age * 7
  }
  
  else if (animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

animal_age(animal = "dog", age = 4)

# animal_age(animal = "hawk", age =  3) #will break! also if you put numbers 
```

### looking at %in%


```{r}
adelie_gentoo <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo")) # look in every row in species colum does it match A or G 
#this will look in any row for Adelie or Gentoo 
```


```{r}
adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))

# when I say look for an exact match with a vector, it is looking for an exact match with the order Adelie and then Gentoo. if not it will be rejected 

# could do this, 

adelie_gentoo_wrong <- penguins %>% 
  filter(species == "Adelie" | species == "Gentoo") # but this will take forever if you have lots of stuff
```

### functions with for loops 

```{r}
df_means <- function(df) {
  
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE) 
    col_name <- colnames(df[i])
    print(paste("The mean value that", col_name, "is", col_mean))
  }
}

df_means(df = mtcars)

# upate the function so that for each iteration is returns the phrase "The mean value that column name 
```

### making a subset of only teh structural size variablles from teh penguins data set for chinstrap

```{r}
chinstrap_sizes <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, flipper_length_mm, bill_depth_mm, body_mass_g)

#now lets use our new function

df_means(df = chinstrap_sizes)
```

### condtion to check teh class of an input 

```{r}
lunchtime <- function(tacos, price_per) {
  
  if (is.numeric(tacos) == FALSE) {
    stop("you must input a number of tacos")
  }
  
  total_price <- tacos * price_per
  return(total_price)
}

lunchtime(tacos = 4, price_per =  2.5)

# lunchtime(tacos = "hi", price_per =  2.5)
```


















